---
title: "cbp_subanalysis"
author: "Christian Pascual"
date: "11/23/2018"
output: github_document
---

```{r setup, message = FALSE, warning = FALSE }
library(caret)
library(tidyverse)
library(leaps)
library(haven)
source("./utils.R")

# Formatting plot output
knitr::opts_chunk$set(
  out.width = "90%"
)

# Set the plot design
theme_set(theme_classic() + 
            theme(legend.position = "bottom", 
                  plot.title = element_text(hjust = 0.5)))

# Raw data
colectomies = read_dta(file = './colectomy_raw_new.dta') 

tidy_colectomies = colectomies %>% 
  select(-starts_with("flg_"), -starts_with("e_")) %>% 
  select_if(unlist(map(., is_mostly_intact), use.names = FALSE)) %>% 
  prettify_names(.) 

tidy_colectomies = tidy_colectomies %>%
  catfactory(.) %>% 
  mutate(any_ssi = (postop_ssi_super + postop_ssi_deep + 	postop_ssi_organspace) >= 1)

small = sample_n(tidy_colectomies, 1000)
```


```{r}
write.csv(small, file = "./sample_colectomies.csv")
```








```{r lib-imports, message = FALSE }
# Our data analysis toolkit
library(tidyverse)

# Formatting plot output
knitr::opts_chunk$set(
  out.width = "90%"
)

# Set the plot design
theme_set(theme_classic() + 
            theme(legend.position = "bottom", 
                  plot.title = element_text(hjust = 0.5)))

colectomies = read.csv(file = './procedure10.csv') 

tidy_colectomies = colectomies %>% 
  select(-contains("flg_"), -contains("e_")) %>%
  select_if(colSums(is.na(.)) < nrow(.) * 0.5)
```

# How are successes associated with ASA level?

Important information

* Covariate: asa_class_id
* Outcome: postop_ssi_super	
* Outcome: postop_ssi_deep	
* Outcome: postop_ssi_organspace

```{r asa-vs-ssi }
asa_analysis = tidy_colectomies %>% 
  select(asa_class_id, postop_ssi_super, postop_ssi_deep, postop_ssi_organspace) %>% 
  mutate(any_ssi = postop_ssi_super + postop_ssi_deep + postop_ssi_organspace,
         had_ssi = any_ssi > 0,
         asa_class_id = as.factor(asa_class_id),
         ssi_type = ifelse(
           postop_ssi_super == 1, "Super",
           ifelse(postop_ssi_deep == 1, "Deep", 
                  ifelse(postop_ssi_organspace == 1, "OrganSpace", "None")
         ))) %>%
  rename(., ASA = asa_class_id) %>% 
  group_by(ASA) %>% 
  summarize(n = n(), 
            n_SSIs = sum(had_ssi),
            percent_SSIs = n_SSIs/n)

knitr::kable(asa_analysis)
```

Now plotting out this analysis:

```{r}
ggplot(data = asa_analysis, aes(x = ASA, y = n, fill = ASA)) +
  geom_bar(stat = "identity")
```

The majority of colectomy surgeries are classified as 2 or 3. 

```{r}
ggplot(data = asa_analysis, aes(x = ASA, y = percent_SSIs, fill = ASA)) +
  geom_bar(stat = "identity")
```

However, we see that there is also an increasing trend in surgical site infections as the ASA severity gets higher. 

```{r distribution-of-ssi-type }
asa_analysis2 = tidy_colectomies %>% 
  select(asa_class_id, postop_ssi_super, postop_ssi_deep, postop_ssi_organspace) %>% 
  mutate(any_ssi = postop_ssi_super + postop_ssi_deep + postop_ssi_organspace,
         had_ssi = any_ssi > 0,
         asa_class_id = as.factor(asa_class_id),
         ssi_type = ifelse(
           postop_ssi_super == 1, "Super",
           ifelse(postop_ssi_deep == 1, "Deep", 
                  ifelse(postop_ssi_organspace == 1, "OrganSpace", "None")
         ))) %>%
  rename(., ASA = asa_class_id) %>% 
  filter(ssi_type != "None")

ggplot(data = asa_analysis2, aes(x = ASA, fill = ssi_type)) +
  geom_bar()
```

Just looking at the instances of SSI, we see that there's not really a discernable pattern in how the types of SSI are distributed by ASA type. 
