---
title: "tiff_subanalysis"
author: "Tiffany Tu"
date: "11/25/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
source("./utils.R")

colectomies = read.csv(file = './procedure10.csv') 

is_mostly_intact = function(col) {
  # Account for literal NAs and blanks
  missings = is.na(col) | col == "" | col == "NA"
  return(sum(missings) <= (3084 / 2))
}

tidy_colectomies = colectomies %>% 
  select(-starts_with("flg_"), -starts_with("e_")) %>% 
  select_if(unlist(map(., is_mostly_intact), use.names = FALSE))

prettify_names(tidy_colectomies)

write.csv(tidy_colectomies, file = "tidycolect.csv")
```

## How are successes distributed by patients with different diseases and prior conditions?

* Success can be measured by Surgical Site Infection (SSI) levels
  - postop_ssi_super	
  - postop_ssi_deep	
  - postop_ssi_organspace

* Pre-operative health conditions acronyms
  - Scd : Specific Carbohydrate Diet
  - Etoh: Alcohol use
  - Copd: Chronic Obstructive Pulmonary Disease
  - Dvt: Deep Vein Thrombosis

There are three levels of severity for variables sleep apnea and diabetes. All variables are converted to binary for this analysis and NA entries are set to 0, indicating that the patient does not have this condition. 

```{r include=FALSE}
healthdisease = tidy_colectomies %>% 
  select(postop_ssi_super, postop_ssi_deep, postop_ssi_organspace, 
         death, ascites, bleeding_disorder, body_weight_loss, 
         chf, copd, chronic_condition, coronary_artery,
         diabetes, dialysis, disseminated_cancer, dvt,
         hypertension, peripheral_vascular_disease,
         pneumonia, preop_sepsis, preop_transfusion,
         sleep_apnea, smoker, etoh, ventilator, beta_blocker, 
         heparinbid, heparintid, indwelling_catheter, scd, 
         preop_sepsis, surgical_approach) %>% 
  mutate(any_ssi = postop_ssi_super + postop_ssi_deep + postop_ssi_organspace,
         diabetes = ifelse(diabetes == 1, 0, 1))
```

```{r warning = FALSE}
dist_healthdisease = healthdisease %>% 
  select(-postop_ssi_super, -postop_ssi_deep, -postop_ssi_organspace, 
         -any_ssi, -death, -surgical_approach) %>% 
  mutate(preop_sepsis = ifelse(preop_sepsis == 1, 0, 1), 
         sleep_apnea = ifelse(sleep_apnea == 0, 0, 1)) %>%
  replace_na(list(dvt = 0, disseminated_cancer = 0, beta_blocker = 0, scd = 0,
                  heparinbid = 0, heparintid = 0, indwelling_catheter = 0)) %>% 
  rownames_to_column %>% 
  gather(condition, value, -rowname) %>% 
  spread(rowname, value) %>% 
  mutate(cases = rowSums(.[2:3085])) %>% 
  transform(condition = reorder(condition, -cases))

ggplot(dist_healthdisease, aes(x = condition, y = cases, fill = condition)) + geom_bar(stat = "identity") + ggtitle("Pre-operative health conditions") + theme(axis.text.x = element_text(angle = 40, hjust = 1), legend.position = "none")
```

```{r}

```

