---
title: "xs_subanalysis"
author: "Xiao Shi"
date: "November 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As discussed in Kaltenbach(2008), Propensity Score Matching (PSM) is designed for the purpose of avoiding comparisons between nonrandomized samples like apples and oranges, for example when assessing two surgeons having different patients. Also, sometimes it is unethical to assign patients to certain treatments in order to get randomized samples. We want to identify, in our hospital rankings case, that the difference in outcomes is attributable to hospitals or patients conditions. The matching scheme provides a way to summarize covariate information about treatment selection into a scalar value (called propensity score) so that matched pairs would have similar scalar value and therefore tend to have similar covariate information. (Kaltenbach, 2008)

Specifically, we identify a specific hospital as a treatment effect. This process will be repeated for every hospital. Then, we model this treatment effect as a function of covariates with logistic regression. The current convention for selecting covariates is to include a potential confounder if unclear of its effect. Therefore, for two patients who are matched because of similar propensity score and one received treatment while the other did not, we can treat them as randomized samples assigned to the treatment/experiment group and control group. Subjects in the matched pairs tend to have similar distributions in covariates (Kaltenbach 2008). We thus have the expression, for each hospital i:

$ln(\frac{PS_i}{1-PS_i}) = \alpha + \beta^T\chi$


PS stands for propensity score. Then we match the propensity score using nearest neighbors with the ratio of 1:5; that is for each treatment patient we match it with the closest 5 patients in the control group. In this way, we get comparable samples adjusted for the apples-to-oranges concerns. Then we do a two sample T-test to calculate the T-scores between these samples and record the T-scores. We then repeat the process for each hospital and get 52 T-scores. These T-scores are ranked to generate the rankings with this method. A positive or high T-score will indicate that the hospital is performing poorly.

The R package ”MatchIt” is used for implementation and the code is attached in the appendix. The tutorial by Simon Ejdemyr on propensity score matching has been a useful reference in implementation.

### Data Cleaning
```{r}
#Hospital Colectomy
library(dummies)
library(haven)
dt = read_dta("colectomy_raw_new.dta")
n=nrow(dt)

knitr::kable(dt$site_cid_160801)

# Cases in each hospital somewhat unbalanced: from 1 to 478.
summary(as.numeric(table(dt$site_cid_160801)))
table(dt$e_gender)
#Men:Woman=1:1
table(dt$e_race)
summary(dt$val_age)

#Site ID as factor
dt$site_cid_160801=substr(dt$site_cid_160801,2,4)
dt$site_cid_160801=as.numeric(dt$site_cid_160801)
dt$site_cid_160801=as.factor(dt$site_cid_160801)

#EDA:
#1.Histogram of cases in each hospital:
library(ggplot2)
ggplot(dt, aes(x = site_cid_160801)) + 
  geom_bar() + 
  labs(x = "Hospital", y = "Number of cases", title = "Histogram of Patients") +
  theme(axis.text.x = element_text(angle = 90))

ggplot(dt, aes(x = as.factor(surgical_approach))) + 
  geom_bar(fill = "blue") + 
  labs(x = "Approach", y = "Number of cases", title = 'Histogram of Surgical Approaches') +
  theme(axis.text.x = element_text(angle = 90))


table(dt$surgical_approach)
#  1    2    3    4    5    6    7   13   14   15   16   17   18 
#5974 1986 1251  891  604   67   67    4   13    6    2    1    2 
# 1 for open, 2 for Laparoscopic, 3 for Laparoscopic, Hand-Assisted
# 4 for Laparoscopic, Converted to Open
# 5 for Robotic
# 6 for Robotic, Converted to Open
# 7 for Laparoscopic, Single Port (SIL)
# 13-18 for Other.

#Remove the columns with mostly missing values: (less than 10000 colectomy cases)
colectomy = dt;
colectomy.var = logical(ncol(dt))
for(i in 1:ncol(dt)){
  complete_case = complete.cases(dt[,i]);
  complete_case = as.numeric(complete_case)
  if(sum(complete_case) > 10000){
    colectomy.var[i]=TRUE
  }
}
table(colectomy.var) #554 variables removed.
```
Reference
1. Lisa Kaltenbach, How To Use Propensity Score Analysis (2008)
2. Simon Ejdemyr, R Tutorial 8: Propensity Score Matching,
http://stanford.edu/ ejdemyr/r-tutorials-archive/tutorial8.html
3. Wikipedia: ASA physical status classification system
